### 2021-01-10 code/data-analysis/factor-analysis.Rmd
Double checked Air quality data processing

---
title: "Air Quality Data from EPA AQS"
author: "Fatemeh Kazemi"
date: "01-01-2021"
output:
  html_document:
    df_print: paged
---

### Last Edit: 12-31-2020

### This program:
  (1) Run factor anlaysis

```{r load packages}
library(tidyverse)
library(here)
library(psych) # for factor analysis

```

```{r Load datasets}
load(here('data','processed', 'pmc-selected.RDa'))
```

```{r - preliminary factor analysis}
# All eligible monitors
fit <- factanal(dt.pmc.sel %>% select(AS:OC), 8, rotation="varimax")
factor.loadings <- rbind(fit$loadings)
save(factor.loadings, file = here('output','table','factor-loadings.RDa'))

factor.scores <- factor.scores(dt.pmc.sel %>% select(AS:OC), fit)
dt.pmc.fac <- cbind(dt.pmc.sel, factor.scores$scores)
save(dt.pmc.fac, file = here('data', 'processed', 'pmc-fac.RDa'))

# Removing EC from factor analysis
fit <- factanal(dt.pmc.sel %>% select(AS:ZN, OC), 8, rotation="varimax")
factor.loadings.ex.ec <- rbind(fit$loadings)
save(factor.loadings.ex.ec, file = here('output','table','factor-loadings-ex-ec.RDa'))

factor.scores <- factor.scores(dt.pmc.sel %>% select(AS:ZN, OC), fit)
dt.pmc.fac.ex.ec <- cbind(dt.pmc.sel, factor.scores$scores)
save(dt.pmc.fac.ex.ec, file = here('data', 'processed', 'pmc-fac-ex-ec.RDa'))
```