### 2021-01-22 code/data-analysis/config-file-generator.Rmd
PMC and CMS Merge Config file generator factor analysis output data cleaning forest plot

---
title: "Config File Generator"
output: html_notebook
---

### This program:
  (1) Generates configuration files for Java for different combinations of 
      i. pollutant
      ii. cause of death
      iii. model

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r List of PM Components}
pmc <- c("Yr_AS","Yr_CA","Yr_CL","Yr_CU","Yr_EC","Yr_FE","Yr_K",
       "Yr_MG","Yr_MN","Yr_NA","Yr_NI","Yr_NO3","Yr_OC","Yr_PB",
       "Yr_PM","Yr_SE","Yr_SI","Yr_SO4","Yr_V","Yr_ZN")
```

```{r List of Factors}
fac <- c("Factor1","Factor2","Factor3","Factor4",
         "Factor5","Factor6","Factor7","Factor8")
```

```{r List of Carbons}
carbon <- c("Yr_EC", "Yr_OC")
```

```{r List of Causes of Death}
cuz <- c("allcuz","nacc","acc","cvd","ihd","chf","cbv","resp","copd","pneu","canc","lungc")
```

```{r Config file - Base Model}
config.base <- c("# data path
data_path=/scratch/fatemehkp/projects/data/analysis/dataset

# loader settings
# separated symbol for the given data
sep=,
# the predictors are included in the model
predictors=exposure
# stratas
stratas=StrID, SiteID
# death
death=no_death_outcome
# life
life=no_enrollee

# optimizer settings
# whether considering ties for death, default false
isTies=false
# if given multiple cores, set isParallelism as true
isParallelism=true
# whether penalizing the weight
l2=false
# if penalizing the weight, set the strength, otherwise it is useless.
l2Strength=1

# the application class name
survival.class=CoxPH
")
```

```{r Config file - SES Adj. Model}
config.ses <- c("# data path
data_path=/scratch/fatemehkp/projects/data/analysis/dataset

# loader settings
# separated symbol for the given data
sep=,
# the predictors are included in the model
predictors=exposure, ses
# stratas
stratas=StrID, SiteID
# death
death=no_death_outcome
# life
life=no_enrollee

# optimizer settings
# whether considering ties for death, default false
isTies=false
# if given multiple cores, set isParallelism as true
isParallelism=true
# whether penalizing the weight
l2=false
# if penalizing the weight, set the strength, otherwise it is useless.
l2Strength=1

# the application class name
survival.class=CoxPH
")
```

```{r base model for components}
dataname <- "ndi-pmc.csv"
expo <- pmc

for (i in seq(1,length(expo),1)){
  filenames <- c( tempfile())
  for( f in  filenames ){
    cat(config.base, file=f)
    x <- readLines(f)
    y <- gsub("dataset", dataname, x)
    z <- gsub("exposure", expo[i], y)
    for (j in seq(1,length(cuz),1)){
      w <- gsub("outcome", cuz[j], z)
      cat(w, file=f, sep="\n")
      k=j+length(cuz)*(i-1)
      writeLines(w, here('config-files','pmc-base', paste(k,'.txt', sep = "")))
    }
  }
  file.remove(f)
  }
```

```{r ses adj. model for components}
dataname <- "ndi-pmc.csv"
expo <- pmc

for (i in seq(1,length(expo),1)){
  filenames <- c( tempfile())
  for( f in  filenames ){
    cat(config.ses, file=f)
    x <- readLines(f)
    y <- gsub("dataset", dataname, x)
    z <- gsub("exposure", expo[i], y)
    for (j in seq(1,length(cuz),1)){
      w <- gsub("outcome", cuz[j], z)
      cat(w, file=f, sep="\n")
      k=j+length(cuz)*(i-1)
      writeLines(w, here('config-files','pmc-ses', paste(k,'.txt', sep = "")))
    }
  }
  file.remove(f)
  }
```

```{r base model for factors}
dataname <- "ndi-fac.csv"
expo <- fac

for (i in seq(1,length(expo),1)){
  filenames <- c( tempfile())
  for( f in  filenames ){
    cat(config.base, file=f)
    x <- readLines(f)
    y <- gsub("dataset", dataname, x)
    z <- gsub("exposure", expo[i], y)
    for (j in seq(1,length(cuz),1)){
      w <- gsub("outcome", cuz[j], z)
      cat(w, file=f, sep="\n")
      k=j+length(cuz)*(i-1)
      writeLines(w, here('config-files','fac-base', paste(k,'.txt', sep = "")))
    }
  }
  file.remove(f)
  }
```

```{r ses adj. model for factors}
dataname <- "ndi-fac.csv"
expo <- fac

for (i in seq(1,length(expo),1)){
  filenames <- c( tempfile())
  for( f in  filenames ){
    cat(config.ses, file=f)
    x <- readLines(f)
    y <- gsub("dataset", dataname, x)
    z <- gsub("exposure", expo[i], y)
    for (j in seq(1,length(cuz),1)){
      w <- gsub("outcome", cuz[j], z)
      cat(w, file=f, sep="\n")
      k=j+length(cuz)*(i-1)
      writeLines(w, here('config-files','fac-ses', paste(k,'.txt', sep = "")))
    }
  }
  file.remove(f)
  }
```

```{r base model for factors excluding EC}
dataname <- "ndi-fac-ex-ec.csv"
expo <- fac

for (i in seq(1,length(expo),1)){
  filenames <- c( tempfile())
  for( f in  filenames ){
    cat(config.base, file=f)
    x <- readLines(f)
    y <- gsub("dataset", dataname, x)
    z <- gsub("exposure", expo[i], y)
    for (j in seq(1,length(cuz),1)){
      w <- gsub("outcome", cuz[j], z)
      cat(w, file=f, sep="\n")
      k=j+length(cuz)*(i-1)
      writeLines(w, here('config-files','fac-ex-ec-base', paste(k,'.txt', sep = "")))
    }
  }
  file.remove(f)
  }
```

```{r ses adj. model for factors excluding EC}
dataname <- "ndi-fac-ex-ec.csv"
expo <- fac

for (i in seq(1,length(expo),1)){
  filenames <- c( tempfile())
  for( f in  filenames ){
    cat(config.ses, file=f)
    x <- readLines(f)
    y <- gsub("dataset", dataname, x)
    z <- gsub("exposure", expo[i], y)
    for (j in seq(1,length(cuz),1)){
      w <- gsub("outcome", cuz[j], z)
      cat(w, file=f, sep="\n")
      k=j+length(cuz)*(i-1)
      writeLines(w, here('config-files','fac-ex-ec-ses', paste(k,'.txt', sep = "")))
    }
  }
  file.remove(f)
  }
```

```{r ses adj. model for Carbons}
dataname <- "ndi-carbon.csv"
expo <- carbon

for (i in seq(1,length(expo),1)){
  filenames <- c( tempfile())
  for( f in  filenames ){
    cat(config.ses, file=f)
    x <- readLines(f)
    y <- gsub("dataset", dataname, x)
    z <- gsub("exposure", expo[i], y)
    for (j in seq(1,length(cuz),1)){
      w <- gsub("outcome", cuz[j], z)
      cat(w, file=f, sep="\n")
      k=j+length(cuz)*(i-1)
      writeLines(w, here('config-files','carbon-ses', paste(k,'.txt', sep = "")))
    }
  }
  file.remove(f)
  }
```

```{r ses adj. model for TOT Carbons}
dataname <- "ndi-carbon-tot.csv"
expo <- carbon

for (i in seq(1,length(expo),1)){
  filenames <- c( tempfile())
  for( f in  filenames ){
    cat(config.ses, file=f)
    x <- readLines(f)
    y <- gsub("dataset", dataname, x)
    z <- gsub("exposure", expo[i], y)
    for (j in seq(1,length(cuz),1)){
      w <- gsub("outcome", cuz[j], z)
      cat(w, file=f, sep="\n")
      k=j+length(cuz)*(i-1)
      writeLines(w, here('config-files','tot-ses', paste(k,'.txt', sep = "")))
    }
  }
  file.remove(f)
  }
```

```{r ses adj. model for TOR Carbons}
dataname <- "ndi-carbon-tor.csv"
expo <- carbon

for (i in seq(1,length(expo),1)){
  filenames <- c( tempfile())
  for( f in  filenames ){
    cat(config.ses, file=f)
    x <- readLines(f)
    y <- gsub("dataset", dataname, x)
    z <- gsub("exposure", expo[i], y)
    for (j in seq(1,length(cuz),1)){
      w <- gsub("outcome", cuz[j], z)
      cat(w, file=f, sep="\n")
      k=j+length(cuz)*(i-1)
      writeLines(w, here('config-files','tor-ses', paste(k,'.txt', sep = "")))
    }
  }
  file.remove(f)
  }
```

```{r under work - parse output}
result = matrix(nrow=1,ncol=4,byrow = TRUE) 
colnames(result) <- c("Expo","Outcome","Beta","SE")

for (i in seq(1,240,1)){
  test <- readLines(here('output','85.txt'))
  important_line2 <- test[min(grep("^death", test))]
  dt <- dt %>% 
    mutate(cuz = gsub(".*death_", "", important_line2))
  important_line <- test[min(grep("^Factor", test))]
  Expo <- gsub("\t.*", "", important_line)
  Coef <- sub(".*\t(.*)\t.*","\\1", important_line)
  SE <- gsub(".*\t", "", important_line)
}

```