### 2021-01-22 code/data-analysis/output-cleaning.Rmd

---
title: "Analysis Output files"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*.

```{r load packages}
library(tidyverse)
library(here)
```

```{r Load data}
load(here('data','processed','pmc-1yr.RDa'))
load(here('data','processed','fac-1yr.RDa'))
load(here('data','processed','fac-ex-ec-1yr.RDa'))
load(here('data','processed','pmc-site-ndi.RDa'))
```

```{r - Parameter/Factor to Index}
### Components
Para.to.Index <- tribble(
  ~Index, ~Parameter,
   "As", "AS",
   "Ca", "CA",
   "Cu", "CU",
   "Cl", "CL",
   "Fe", "FE",
   "Pb", "PB",
   "Mn", "MN",
   "Ni", "NI",
   "Mg", "MG",
   "Se", "SE",
   "V", "V",
   "Si", "SI",
   "Zn", "ZN",
   "K", "K",
   "Na", "NA",
   "NO3", "NO3",
   "SO4", "SO4",
   "OC", "OC",
   "EC", "EC",
   "PM2.5", "PM"
   ) %>% 
  mutate(Index = as.factor(Index),
         Parameter = as.factor(Parameter))

### Factors
Factor.to.Index <- tribble(
  ~Index, ~Parameter,
   "Traffic", "Factor1",
   "Soil", "Factor2",
   "Metals", "Factor3",
   "Biomass", "Factor4",
   "Salt", "Factor5",
   "Oil", "Factor6",
   "Coal", "Factor7",
   "Steel", "Factor8"
   ) %>% 
  mutate(Index = as.factor(Index),
         Parameter = as.factor(Parameter))

### Factors excluding EC
Factor.ex.ec.to.Index <- tribble(
  ~Index, ~Parameter,
   "Soil", "Factor1",
   "Metals", "Factor2",
   "Traffic", "Factor3",
   "Biomass", "Factor4",
   "Coal", "Factor5",
   "Salt", "Factor6",
   "Oil", "Factor7",
   "Steel", "Factor8"
   ) %>% 
  mutate(Index = as.factor(Index),
         Parameter = as.factor(Parameter))
```

```{r Outcome to Cause }
Outcome.to.Cause <- tribble(
  ~Cause, ~Outcome,
  "All-Cause", "allcuz",
  "Non-Accidental", "nacc",
  "Accidental", "acc",
  "Cardiovascular", "cvd",
  "IHD", "ihd",
  "CHF", "chf",
  "CBV", "cbv",
  "Respiratory", "resp",
  "COPD", "copd",
  "Pneumonia", "pneu",
  "Cancer", "canc",
  "Lung Cancer", "lungc"
)
```

```{r calculating IQRS}
### Components
pmc.1yr.iqr <- dt.pmc.site.ndi %>% 
  select(Site.ID) %>% 
  inner_join(dt.pmc.1yr) %>% 
  summarise_at(vars(AS:ZN), ~IQR(.x)) %>% 
  pivot_longer(cols = AS:ZN,
               names_to = "Parameter",
               values_to = "IQR") %>% 
  mutate(Parameter = as.factor(Parameter))

save(pmc.1yr.iqr, file = here('data','processed','pmc-1yr-iqr.RDa'))

### Factors
fac.1yr.iqr <- dt.pmc.site.ndi %>% 
  select(Site.ID) %>% 
  inner_join(dt.fac.1yr) %>% 
  summarise_at(vars(Factor1:Factor8), ~IQR(.x)) %>% 
  pivot_longer(cols = Factor1:Factor8,
               names_to = "Parameter",
               values_to = "IQR")%>% 
  mutate(Parameter = as.factor(Parameter))

save(fac.1yr.iqr, file = here('data','processed','fac-1yr-iqr.RDa'))

### Factors excluding EC
fac.ex.ec.1yr.iqr <- dt.pmc.site.ndi %>% 
  select(Site.ID) %>% 
  inner_join(dt.fac.ex.ec.1yr) %>% 
  summarise_at(vars(Factor1:Factor8), ~IQR(.x)) %>% 
  pivot_longer(cols = Factor1:Factor8,
               names_to = "Parameter",
               values_to = "IQR")%>% 
  mutate(Parameter = as.factor(Parameter))

save(fac.ex.ec.1yr.iqr, file = here('data','processed','fac-ex-ec-1yr-iqr.RDa'))

### Carbons
dt.carbon <- dt.pmc.1yr %>% 
  select(Site.ID, Year, Month, EC, OC)

dt.site <- dt.pmc.site.ndi %>% 
  filter(!Method == "EC.TOT-OC.TOR") %>% 
  select(Site.ID, Method)

carbon.1yr.iqr <- dt.site %>% 
  select(Site.ID) %>% 
  inner_join(dt.carbon) %>% 
  summarise_at(vars(EC:OC), ~IQR(.x)) %>% 
  pivot_longer(cols = EC:OC,
               names_to = "Parameter",
               values_to = "IQR") %>% 
  mutate(Parameter = as.factor(Parameter))

save(carbon.1yr.iqr, file = here('data','processed','carbon-1yr-iqr.RDa'))

carbon.method.1yr.iqr <- dt.site %>% 
  inner_join(dt.carbon) %>% 
  group_by(Method) %>% 
  summarise_at(vars(EC:OC), ~IQR(.x)) %>% 
  pivot_longer(cols = EC:OC,
               names_to = "Parameter",
               values_to = "IQR") %>% 
  mutate(Parameter = as.factor(Parameter)) %>% 
  arrange(Parameter)

save(carbon.method.1yr.iqr, file = here('data','processed','carbon-method-1yr-iqr.RDa'))
```

```{r Output PMC Base}
out.pmc.base <- read.csv(here('output','analysis','pmc_base.csv')) %>% 
  mutate(Parameter = gsub(".*Yr_", "", Predictor),
         Parameter = as.factor(Parameter)) %>% 
  merge(pmc.1yr.iqr) %>% 
  merge(Para.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, HR, HR.L, HR.U)

save(out.pmc.base, file = here('output','analysis','out-pmc-base.RDa'))
```

```{r Output PMC SES-Adj}
out.pmc.ses <- read.csv(here('output','analysis','pmc_ses.csv')) %>% 
  mutate(Parameter = gsub(".*Yr_", "", Predictor),
         Parameter = as.factor(Parameter)) %>% 
  merge(pmc.1yr.iqr) %>% 
  merge(Para.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, HR, HR.L, HR.U)

save(out.pmc.ses, file = here('output','analysis','out-pmc-ses.RDa'))
```

```{r Output Factors Base}
out.fac.base <- read.csv(here('output','analysis','fac_base.csv')) %>% 
  rename(Parameter = Predictor) %>% 
  merge(fac.1yr.iqr) %>% 
  merge(Factor.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, HR, HR.L, HR.U)

save(out.fac.base, file = here('output','analysis','out-fac-base.RDa'))
```

```{r Output Factors SES-Adj}
out.fac.ses <- read.csv(here('output','analysis','fac_ses.csv')) %>% 
  rename(Parameter = Predictor) %>% 
  merge(fac.1yr.iqr) %>% 
  merge(Factor.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, HR, HR.L, HR.U)

save(out.fac.ses, file = here('output','analysis','out-fac-ses.RDa'))
```

```{r Output Factors Ex. EC Base}
out.fac.ex.ec.base <- read.csv(here('output','analysis','fac_ex_ec_base.csv')) %>% 
  rename(Parameter = Predictor) %>% 
  merge(fac.ex.ec.1yr.iqr) %>% 
  merge(Factor.ex.ec.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, HR, HR.L, HR.U)

save(out.fac.ex.ec.base, file = here('output','analysis','out-fac-ex-ec-base.RDa'))
```

```{r Output Factors Ex. EC SES-Adj}
out.fac.ex.ec.ses <- read.csv(here('output','analysis','fac_ex_ec_ses.csv')) %>% 
  rename(Parameter = Predictor) %>% 
  merge(fac.ex.ec.1yr.iqr) %>% 
  merge(Factor.ex.ec.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, HR, HR.L, HR.U)

save(out.fac.ex.ec.ses, file = here('output','analysis','out-fac-ex-ec-ses.RDa'))
```

```{r Output Carbon SES-Adj}
out.carbon.ses <- read.csv(here('output','analysis','carbon_ses.csv')) %>% 
  mutate(Parameter = gsub(".*Yr_", "", Predictor),
         Parameter = as.factor(Parameter)) %>% 
  merge(carbon.1yr.iqr) %>% 
  merge(Para.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, HR, HR.L, HR.U)

save(out.carbon.ses, file = here('output','analysis','out-carbon-ses.RDa'))
```

```{r Output Carbon TOT SES-Adj}
out.carbon.tot.ses <- read.csv(here('output','analysis','tot_ses.csv')) %>% 
  mutate(Parameter = gsub(".*Yr_", "", Predictor),
         Parameter = as.factor(Parameter)) %>% 
  merge((carbon.method.1yr.iqr %>% filter(Method == "TOT"))) %>% 
  merge(Para.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, HR, HR.L, HR.U)

save(out.carbon.tot.ses, file = here('output','analysis','out-carbon-tot-ses.RDa'))
```

```{r Output Carbon TOR SES-Adj}
out.carbon.tor.ses <- read.csv(here('output','analysis','tot_ses.csv')) %>% 
  mutate(Parameter = gsub(".*Yr_", "", Predictor),
         Parameter = as.factor(Parameter)) %>% 
  merge((carbon.method.1yr.iqr %>% filter(Method == "TOR"))) %>% 
  merge(Para.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, HR, HR.L, HR.U)

save(out.carbon.tor.ses, file = here('output','analysis','out-carbon-tor-ses.RDa'))
```