### 2021-01-22 code/data-processors/cms-data-processor.Rmd

---
title: "Air Quality and CMS Data"
author: "Fatemeh Kazemi"
date: "1/18/2021"
output: html_document
---

### This program:
  (1) Merges CMS(ndi) data with pmc and ses data

### Note:
  1. In data merged with CMS, bufferzone is 12-mile unless specified
  2. Exposure window is one year unless specified

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages}
library(tidyverse)
library(here)
```

```{r load data}
dirUSSpatial <- 'C:\\Users\\fkazem01\\Box\\Projects\\USA Spatial\\data\\processed\\'
load(paste0(dirUSSpatial,'ses-site-bz12.RDa'))
load(here('data','processed','pmc-1yr.RDa'))
load(here('data','processed','fac-1yr.RDa'))
load(here('data','processed','fac-ex-ec-1yr.RDa'))
load(here('data','processed','master-ndi-pmc.RDa'))
load(here('data','processed','pmc-site-ndi.RDa'))
```

```{r Import and save CMS(/ndi) data}
dt.master.ndi.pmc <- read.csv(here('data', 'processed','master_ndi_pmc.csv')) %>% 
  rename(Site.ID = Site_ID,
         Year = year,
         Month = MONTH,
         age = ENROLLEE_AGE)
save(dt.master.ndi.pmc, file = here('data', 'processed', 'master-ndi-pmc.RDa'))

dt.master.ndi.pmc.race <- read.csv(here('data', 'processed','master_ndi_pmc_race.csv')) %>%
  rename(Site.ID = Site_ID,
         Year = year,
         Month = MONTH,
         age = ENROLLEE_AGE)
save(dt.master.ndi.pmc.race, file = here('data', 'processed', 'master-ndi-pmc-race.RDa'))
```

```{r PMC}
dt.pmc <- dt.pmc.1yr %>% 
  rename_at(vars(AS:ZN), ~paste("Yr", .x, sep = "_"))

dt.ses <- dt.ses.site.bz12 %>% 
  mutate(ses = SES.Zip.Norm) %>% 
  select(Site.ID, Year, ses)

dt.ndi.pmc <- dt.master.ndi.pmc %>% 
  mutate(StrID = paste(age, sex, race)) %>% 
  select(Site.ID, Year, Month, StrID, no_enrollee:no_death_kidn) %>%  
  inner_join(dt.pmc) %>% 
  inner_join(dt.ses) %>% 
  rename(SiteID = Site.ID)

write.csv(dt.ndi.pmc, file = here('data', 'processed', 'ndi-pmc.csv'),
          row.names = F)

dt.ndi.pmc.race <- dt.master.ndi.pmc.race %>% 
  mutate(StrID = paste(age, sex, race)) %>% 
  select(Site.ID, Year, Month, StrID, no_enrollee:no_death_kidn) %>% 
  inner_join(dt.pmc) %>% 
  inner_join(dt.ses) %>% 
  rename(SiteID = Site.ID)

write.csv(dt.ndi.pmc.race, file = here('data', 'processed', 'ndi-pmc-race.csv'),
          row.names = F)
```

```{r Factors}
dt.ses <- dt.ses.site.bz12 %>% 
  mutate(ses = SES.Zip.Norm) %>% 
  select(Site.ID, Year, ses)

dt.ndi.fac <- dt.master.ndi.pmc %>% 
  mutate(StrID = paste(age, sex, race)) %>% 
  select(Site.ID, Year, Month, StrID, no_enrollee:no_death_kidn) %>%  
  inner_join(dt.fac.1yr) %>% 
  inner_join(dt.ses) %>% 
  rename(SiteID = Site.ID)

write.csv(dt.ndi.fac, file = here('data', 'processed', 'ndi-fac.csv'),
          row.names = F)

dt.ndi.fac.race <- dt.master.ndi.fac.pmc %>% 
  mutate(StrID = paste(age, sex, race)) %>% 
  select(Site.ID, Year, Month, StrID, no_enrollee:no_death_kidn) %>% 
  inner_join(dt.fac.1yr) %>% 
  inner_join(dt.ses) %>% 
  rename(SiteID = Site.ID)

write.csv(dt.ndi.fac.race, file = here('data', 'processed', 'ndi-fac-race.csv'),
          row.names = F)
```

```{r Factors Excluding EC}
dt.ses <- dt.ses.site.bz12 %>% 
  mutate(ses = SES.Zip.Norm) %>% 
  select(Site.ID, Year, ses)

dt.ndi.fac.ex.ec <- dt.master.ndi.pmc %>% 
  mutate(StrID = paste(age, sex, race)) %>% 
  select(Site.ID, Year, Month, StrID, no_enrollee:no_death_kidn) %>%  
  inner_join(dt.fac.ex.ec.1yr) %>% 
  inner_join(dt.ses) %>% 
  rename(SiteID = Site.ID)

write.csv(dt.ndi.fac.ex.ec, file = here('data', 'processed', 'ndi-fac-ex-ec.csv'),
          row.names = F)

dt.ndi.fac.ex.ec.race <- dt.master.ndi.pmc %>% 
  mutate(StrID = paste(age, sex, race)) %>% 
  select(Site.ID, Year, Month, StrID, no_enrollee:no_death_kidn) %>% 
  inner_join(dt.fac.ex.ec.1yr) %>% 
  inner_join(dt.ses) %>% 
  rename(SiteID = Site.ID)

write.csv(dt.ndi.fac.race, file = here('data', 'processed', 'ndi-fac-ex-ec-race.csv'),
          row.names = F)
```

```{r Carbons}
dt.carbon <- dt.pmc.1yr %>% 
  select(Site.ID, Year, Month, EC, OC) %>% 
  rename_at(vars(EC, OC), ~paste("Yr", .x, sep = "_"))

dt.site <- dt.pmc.site.ndi %>% 
  filter(!Method == "EC.TOT-OC.TOR") %>% 
  select(Site.ID, Method)

dt.ses <- dt.ses.site.bz12 %>% 
  mutate(ses = SES.Zip.Norm) %>% 
  select(Site.ID, Year, ses)

dt.ndi.carbon <- dt.master.ndi.pmc %>% 
  mutate(StrID = paste(age, sex, race)) %>% 
  select(Site.ID, Year, Month, StrID, no_enrollee:no_death_lungc,
         -no_death_uri, -no_death_ards) %>%  
  inner_join(dt.carbon) %>% 
  inner_join(dt.site) %>% 
  inner_join(dt.ses) %>% 
  rename(SiteID = Site.ID) %>% 
  select(-Method)

write.csv(dt.ndi.carbon, file = here('data', 'processed', 'ndi-carbon.csv'),
          row.names = F)

dt.ndi.carbon.tot <- dt.master.ndi.pmc %>% 
  mutate(StrID = paste(age, sex, race)) %>% 
  select(Site.ID, Year, Month, StrID, no_enrollee:no_death_lungc,
         -no_death_uri, -no_death_ards) %>%  
  inner_join(dt.carbon) %>% 
  inner_join(dt.site) %>% 
  inner_join(dt.ses) %>% 
  rename(SiteID = Site.ID) %>% 
  filter(Method == "TOT") %>% 
  select(-Method)

write.csv(dt.ndi.carbon.tot, file = here('data', 'processed', 'ndi-carbon-tot.csv'),
          row.names = F)

dt.ndi.carbon.tor <- dt.master.ndi.pmc %>% 
  mutate(StrID = paste(age, sex, race)) %>% 
  select(Site.ID, Year, Month, StrID, no_enrollee:no_death_lungc,
         -no_death_uri, -no_death_ards) %>%  
  inner_join(dt.carbon) %>% 
  inner_join(dt.site) %>% 
  inner_join(dt.ses) %>% 
  rename(SiteID = Site.ID) %>% 
  filter(Method == "TOR") %>% 
  select(-Method)

write.csv(dt.ndi.carbon.tor, file = here('data', 'processed', 'ndi-carbon-tor.csv'),
          row.names = F)
```