### 2021-01-24 code/data-analysis/factor-analysis.Rmd

---
title: "Air Quality Data from EPA AQS"
author: "Fatemeh Kazemi"
date: "01-01-2021"
output:
  html_document:
    df_print: paged
---

### Last Edit: 12-31-2020

### This program:
  (1) Run factor analysis

```{r load packages}
library(tidyverse)
library(here)
library(psych) # for factor analysis
```

```{r Load datasets}
load(here('data','processed', 'pmc-daily.RDa'))
```

```{r Call functions}
source(here('code','data-processors','air-data-smooth-mvavg-function.R'))
```

```{r - preliminary factor analysis}
# All eligible monitors
fit <- factanal(dt.pmc.daily %>% select(AS:OC), 8, rotation="varimax")
factor.loadings <- rbind(fit$loadings)
save(factor.loadings, file = here('output','table','factor-loadings.RDa'))
#Traffic Soil Metals Biomass Salt Oil Coal Steel

factor.scores <- factor.scores(dt.pmc.daily %>% select(AS:OC), fit)
dt.fac.daily <- cbind(dt.pmc.daily %>% select(Site.ID, Date), factor.scores$scores) 
save(dt.fac.daily, file = here('data', 'processed', 'fac-daily.RDa'))
```

```{r Removing EC from factor analysis}
fit <- factanal(dt.pmc.daily %>% select(AS:ZN, OC), 8, rotation="varimax")
factor.loadings.ex.ec <- rbind(fit$loadings)
save(factor.loadings.ex.ec, file = here('output','table','factor-loadings-ex-ec.RDa'))
# Soil Metals Traffic Biomass Coal Salt Oil Steel

factor.scores <- factor.scores(dt.pmc.daily %>% select(AS:ZN, OC), fit)
dt.fac.daily.ex.ec <- cbind(dt.pmc.daily %>% select(Site.ID, Date), factor.scores$scores)
save(dt.fac.daily.ex.ec, file = here('data', 'processed', 'fac-daily-ex-ec.RDa'))
```

```{r Smoothing Data & Caluculating 1-yr mvavg for each month}
dt.air <- dt.fac.daily %>% 
  pivot_longer(cols = Factor1:Factor8,
               names_to = "Parameter.Symbol",
               values_to = "Value") %>% 
  arrange(Site.ID, Parameter.Symbol, Date)

site.para <- dt.air %>% 
  distinct(Site.ID, Parameter.Symbol)

#using smoothed.yr function
smooth.out <- mapply(Smoothed.Year, site.para$Site.ID, site.para$Parameter.Symbol)
para.monthly <- data.frame(Site.ID=unlist(smooth.out[1,]),
                           Parameter.Symbol=unlist(smooth.out[2,]),
                           Date=unlist(smooth.out[3,]), 
                           para.1yr=unlist(smooth.out[4,]))

dt.fac.1yr <- para.monthly %>% 
  mutate(Year = floor(Date -0.01),
         Month = round((Date - Year)*12)) %>% 
  select(-Date) %>% 
  pivot_wider(names_sort = T,
              names_from = c(Parameter.Symbol),
              values_from = c(para.1yr)) %>% 
  arrange(Site.ID, Year, Month)

save(dt.fac.1yr, file = here('data', 'processed', 'fac-1yr.RDa'))
write.csv(dt.fac.1yr, file = here('data', 'processed', 'fac-1yr.csv'), row.names = F)
```

```{r Smoothing Data & Caluculating 1-yr mvavg for each month}
dt.air <- dt.fac.daily.ex.ec %>% 
  pivot_longer(cols = Factor1:Factor8,
               names_to = "Parameter.Symbol",
               values_to = "Value") %>% 
  arrange(Site.ID, Parameter.Symbol, Date)

site.para <- dt.air %>% 
  distinct(Site.ID, Parameter.Symbol)

#using smoothed.yr function
smooth.out <- mapply(Smoothed.Year, site.para$Site.ID, site.para$Parameter.Symbol)
para.monthly <- data.frame(Site.ID=unlist(smooth.out[1,]),
                           Parameter.Symbol=unlist(smooth.out[2,]),
                           Date=unlist(smooth.out[3,]), 
                           para.1yr=unlist(smooth.out[4,]))

dt.fac.ex.ec.1yr <- para.monthly %>% 
  mutate(Year = floor(Date -0.01),
         Month = round((Date - Year)*12)) %>% 
  select(-Date) %>% 
  pivot_wider(names_sort = T,
              names_from = c(Parameter.Symbol),
              values_from = c(para.1yr)) %>% 
  arrange(Site.ID, Year, Month)

save(dt.fac.ex.ec.1yr, file = here('data', 'processed', 'fac-ex-ec-1yr.RDa'))
write.csv(dt.fac.ex.ec.1yr, file = here('data', 'processed', 'fac-ex-ec-1yr.csv'), row.names = F)
```

```{r residuals of PM vs. Factors}
dt.pm.daily <- dt.pmc.daily %>% 
  select(Site.ID, Date, PM)

dt.fac.ex.ec.residual <- dt.fac.daily.ex.ec %>% 
  merge(dt.pm.daily) %>% 
  group_by(Site.ID) %>% 
  nest() %>% 
  mutate(PM_Factor1 = map(map(data, ~lm(PM ~ Factor1, data = .)), residuals),
         PM_Factor2 = map(map(data, ~lm(PM ~ Factor2, data = .)), residuals),
         PM_Factor3 = map(map(data, ~lm(PM ~ Factor3, data = .)), residuals),
         PM_Factor4 = map(map(data, ~lm(PM ~ Factor4, data = .)), residuals),
         PM_Factor5 = map(map(data, ~lm(PM ~ Factor5, data = .)), residuals),
         PM_Factor6  = map(map(data, ~lm(PM ~ Factor6,  data = .)), residuals),
         PM_Factor7 = map(map(data, ~lm(PM ~ Factor7, data = .)), residuals),
         PM_Factor8 = map(map(data, ~lm(PM ~ Factor8, data = .)), residuals)
         ) %>% 
  unnest(c(data, PM_Factor1:PM_Factor8)) %>% 
  select(-(Factor1:Factor8), -PM)

save(dt.fac.ex.ec.residual, file = here('data', 'processed', 'fac-ex-ec-residual.RDa'))
```

```{r Smoothing Data & Caluculating 1-yr mvavg for each month}
dt.air <- dt.fac.ex.ec.residual %>% 
  pivot_longer(cols = PM_Factor1:PM_Factor8,
               names_to = "Parameter.Symbol",
               values_to = "Value") %>% 
  arrange(Site.ID, Parameter.Symbol, Date)

site.para <- dt.air %>% 
  distinct(Site.ID, Parameter.Symbol)

#using smoothed.yr function
smooth.out <- mapply(Smoothed.Year, site.para$Site.ID, site.para$Parameter.Symbol)
para.monthly <- data.frame(Site.ID=unlist(smooth.out[1,]),
                           Parameter.Symbol=unlist(smooth.out[2,]),
                           Date=unlist(smooth.out[3,]), 
                           para.1yr=unlist(smooth.out[4,]))

dt.fac.ex.ec.residual.1yr <- para.monthly %>% 
  mutate(Year = floor(Date -0.01),
         Month = round((Date - Year)*12)) %>% 
  select(-Date) %>% 
  pivot_wider(names_sort = T,
              names_from = c(Parameter.Symbol),
              values_from = c(para.1yr)) %>% 
  arrange(Site.ID, Year, Month)

save(dt.fac.ex.ec.residual.1yr, file = here('data', 'processed', 'fac-ex-ec-residual-1yr.RDa'))

```