### 2021-01-29 code/data-analysis/output-cleaning.Rmd

---
title: "Analysis Output files"
---

```{r load packages}
library(tidyverse)
library(here)
```

```{r Load data}
load(here('data','processed','pmc-1yr.RDa'))
load(here('data','processed','fac-1yr.RDa'))
load(here('data','processed','fac-ex-ec-1yr.RDa'))
load(here('data','processed','pmc-site-ndi.RDa'))
load(here('data','processed','pmc-sub-1yr-5yr.RDa'))
load(here('data','processed','pmc-sub-2yr-5yr.RDa'))
load(here('data','processed','pmc-sub-3yr-5yr.RDa'))
load(here('data','processed','pmc-sub-4yr-5yr.RDa'))
load(here('data','processed','pmc-sub-5yr.RDa'))
```

```{r - Parameter/Factor to Index}
### Components
Para.to.Index <- tribble(
  ~Index, ~Parameter,
   "As", "AS",
   "Ca", "CA",
   "Cu", "CU",
   "Cl", "CL",
   "Fe", "FE",
   "Pb", "PB",
   "Mn", "MN",
   "Ni", "NI",
   "Mg", "MG",
   "Se", "SE",
   "V", "V",
   "Si", "SI",
   "Zn", "ZN",
   "K", "K",
   "Na", "NA",
   "NO3", "NO3",
   "SO4", "SO4",
   "OC", "OC",
   "EC", "EC",
   "PM2.5", "PM"
   ) %>% 
  mutate(Index = as.factor(Index),
         Parameter = as.factor(Parameter))

### Factors
Factor.to.Index <- tribble(
  ~Index, ~Parameter,
   "Traffic", "Factor1",
   "Soil", "Factor2",
   "Metals", "Factor3",
   "Biomass", "Factor4",
   "Salt", "Factor5",
   "Oil", "Factor6",
   "Coal", "Factor7",
   "Steel", "Factor8"
   ) %>% 
  mutate(Index = as.factor(Index),
         Parameter = as.factor(Parameter))

### Factors excluding EC
Factor.ex.ec.to.Index <- tribble(
  ~Index, ~Parameter,
   "Soil", "Factor1",
   "Metals", "Factor2",
   "Traffic", "Factor3",
   "Biomass", "Factor4",
   "Coal", "Factor5",
   "Salt", "Factor6",
   "Oil", "Factor7",
   "Steel", "Factor8"
   ) %>% 
  mutate(Index = as.factor(Index),
         Parameter = as.factor(Parameter))
```

```{r Outcome to Cause }
Outcome.to.Cause <- tribble(
  ~Cause, ~Outcome,
  "All-Cause", "allcuz",
  "Non-Accidental", "nacc",
  "Accidental", "acc",
  "Cardiovascular", "cvd",
  "IHD", "ihd",
  "CHF", "chf",
  "CBV", "cbv",
  "Respiratory", "resp",
  "COPD", "copd",
  "Pneumonia", "pneu",
  "Cancer", "canc",
  "Lung Cancer", "lungc"
)
```

```{r calculating IQRS}
### Components
pmc.1yr.iqr <- dt.pmc.site.ndi %>% 
  select(Site.ID) %>% 
  inner_join(dt.pmc.1yr) %>% 
  summarise_at(vars(AS:ZN), ~IQR(.x)) %>% 
  pivot_longer(cols = AS:ZN,
               names_to = "Parameter",
               values_to = "IQR") %>% 
  mutate(Parameter = as.factor(Parameter))

### Factors
fac.1yr.iqr <- dt.pmc.site.ndi %>% 
  select(Site.ID) %>% 
  inner_join(dt.fac.1yr) %>% 
  summarise_at(vars(Factor1:Factor8), ~IQR(.x)) %>% 
  pivot_longer(cols = Factor1:Factor8,
               names_to = "Parameter",
               values_to = "IQR")%>% 
  mutate(Parameter = as.factor(Parameter))

### Factors excluding EC
fac.ex.ec.1yr.iqr <- dt.pmc.site.ndi %>% 
  select(Site.ID) %>% 
  inner_join(dt.fac.ex.ec.1yr) %>% 
  summarise_at(vars(Factor1:Factor8), ~IQR(.x)) %>% 
  pivot_longer(cols = Factor1:Factor8,
               names_to = "Parameter",
               values_to = "IQR")%>% 
  mutate(Parameter = as.factor(Parameter))

### Carbons
dt.carbon <- dt.pmc.1yr %>% 
  select(Site.ID, Year, Month, EC, OC)

dt.site <- dt.pmc.site.ndi %>% 
  filter(!Method == "EC.TOT-OC.TOR") %>% 
  select(Site.ID, Method)

carbon.1yr.iqr <- dt.site %>% 
  select(Site.ID) %>% 
  inner_join(dt.carbon) %>% 
  summarise_at(vars(EC:OC), ~IQR(.x)) %>% 
  pivot_longer(cols = EC:OC,
               names_to = "Parameter",
               values_to = "IQR") %>% 
  mutate(Parameter = as.factor(Parameter))

carbon.method.1yr.iqr <- dt.site %>% 
  inner_join(dt.carbon) %>% 
  group_by(Method) %>% 
  summarise_at(vars(EC:OC), ~IQR(.x)) %>% 
  pivot_longer(cols = EC:OC,
               names_to = "Parameter",
               values_to = "IQR") %>% 
  mutate(Parameter = as.factor(Parameter)) %>% 
  arrange(Parameter)

### Components - urbanicity
pmc.1yr.locset.iqr <- dt.pmc.site.ndi %>% 
   mutate(Loc.Set = ifelse(!Location.Setting == "RURAL", "urban","rural")) %>% 
  select(Site.ID, Loc.Set) %>% 
  inner_join(dt.pmc.1yr) %>% 
  group_by(Loc.Set) %>% 
  summarise_at(vars(AS:ZN), ~IQR(.x)) %>% 
  pivot_longer(cols = AS:ZN,
               names_to = "Parameter",
               values_to = "IQR") %>% 
  mutate(Parameter = as.factor(Parameter))

### Components - region
pmc.1yr.region.iqr <- dt.pmc.site.ndi %>% 
  select(Site.ID, Region.IV) %>% 
  inner_join(dt.pmc.1yr) %>% 
  group_by(Region.IV) %>% 
  summarise_at(vars(AS:ZN), ~IQR(.x)) %>% 
  pivot_longer(cols = AS:ZN,
               names_to = "Parameter",
               values_to = "IQR") %>% 
  mutate(Parameter = as.factor(Parameter))

### Components - 5yr
pmc.5yr.iqr <- dt.pmc.5yr %>% 
  summarise_at(vars(AS:ZN), ~IQR(.x)) %>% 
  pivot_longer(cols = AS:ZN,
               names_to = "Parameter",
               values_to = "IQR") %>% 
  mutate(Parameter = as.factor(Parameter))

### Components - 4yr
pmc.4yr.5yr.iqr <- dt.pmc.4yr.5yr %>% 
  summarise_at(vars(AS:ZN), ~IQR(.x)) %>% 
  pivot_longer(cols = AS:ZN,
               names_to = "Parameter",
               values_to = "IQR") %>% 
  mutate(Parameter = as.factor(Parameter))

### Components - 3yr
pmc.3yr.5yr.iqr <- dt.pmc.3yr.5yr %>% 
  summarise_at(vars(AS:ZN), ~IQR(.x)) %>% 
  pivot_longer(cols = AS:ZN,
               names_to = "Parameter",
               values_to = "IQR") %>% 
  mutate(Parameter = as.factor(Parameter))

### Components - 2yr
pmc.2yr.5yr.iqr <- dt.pmc.2yr.5yr %>% 
  summarise_at(vars(AS:ZN), ~IQR(.x)) %>% 
  pivot_longer(cols = AS:ZN,
               names_to = "Parameter",
               values_to = "IQR") %>% 
  mutate(Parameter = as.factor(Parameter))

### Components - 1yr
pmc.1yr.5yr.iqr <- dt.pmc.1yr.5yr %>% 
  summarise_at(vars(AS:ZN), ~IQR(.x)) %>% 
  pivot_longer(cols = AS:ZN,
               names_to = "Parameter",
               values_to = "IQR") %>% 
  mutate(Parameter = as.factor(Parameter))
```

```{r Output PMC Base}
out.pmc.base <- read.csv(here('output','analysis','pmc_base.csv')) %>% 
  mutate(Parameter = gsub(".*Yr_", "", Predictor),
         Parameter = as.factor(Parameter)) %>% 
  merge(pmc.1yr.iqr) %>% 
  merge(Para.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, HR, HR.L, HR.U)

save(out.pmc.base, file = here('output','analysis','out-pmc-base.RDa'))
```

```{r Output PMC SES-Adj}
out.pmc.ses <- read.csv(here('output','analysis','pmc_ses.csv')) %>% 
  mutate(Parameter = gsub(".*Yr_", "", Predictor),
         Parameter = as.factor(Parameter)) %>% 
  merge(pmc.1yr.iqr) %>% 
  merge(Para.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, HR, HR.L, HR.U)

save(out.pmc.ses, file = here('output','analysis','out-pmc-ses.RDa'))
```

```{r Output Factors Base}
out.fac.base <- read.csv(here('output','analysis','fac_base.csv')) %>% 
  rename(Parameter = Predictor) %>% 
  merge(fac.1yr.iqr) %>% 
  merge(Factor.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, HR, HR.L, HR.U)

save(out.fac.base, file = here('output','analysis','out-fac-base.RDa'))
```

```{r Output Factors SES-Adj}
out.fac.ses <- read.csv(here('output','analysis','fac_ses.csv')) %>% 
  rename(Parameter = Predictor) %>% 
  merge(fac.1yr.iqr) %>% 
  merge(Factor.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, HR, HR.L, HR.U)

save(out.fac.ses, file = here('output','analysis','out-fac-ses.RDa'))
```

```{r Output Factors Ex. EC Base}
out.fac.ex.ec.base <- read.csv(here('output','analysis','fac_ex_ec_base.csv')) %>% 
  rename(Parameter = Predictor) %>% 
  merge(fac.ex.ec.1yr.iqr) %>% 
  merge(Factor.ex.ec.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, HR, HR.L, HR.U)

save(out.fac.ex.ec.base, file = here('output','analysis','out-fac-ex-ec-base.RDa'))
```

```{r Output Factors Ex. EC SES-Adj}
out.fac.ex.ec.ses <- read.csv(here('output','analysis','fac_ex_ec_ses.csv')) %>% 
  rename(Parameter = Predictor) %>% 
  merge(fac.ex.ec.1yr.iqr) %>% 
  merge(Factor.ex.ec.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, HR, HR.L, HR.U)

save(out.fac.ex.ec.ses, file = here('output','analysis','out-fac-ex-ec-ses.RDa'))
```

```{r Output Carbon SES-Adj}
out.carbon.ses <- read.csv(here('output','analysis','carbon_ses.csv')) %>% 
  mutate(Parameter = gsub(".*Yr_", "", Predictor),
         Parameter = as.factor(Parameter)) %>% 
  merge(carbon.1yr.iqr) %>% 
  merge(Para.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, HR, HR.L, HR.U)

save(out.carbon.ses, file = here('output','analysis','out-carbon-ses.RDa'))
```

```{r Output Carbon TOT SES-Adj}
out.carbon.tot.ses <- read.csv(here('output','analysis','tot_ses.csv')) %>% 
  mutate(Parameter = gsub(".*Yr_", "", Predictor),
         Parameter = as.factor(Parameter)) %>% 
  merge((carbon.method.1yr.iqr %>% filter(Method == "TOT"))) %>% 
  merge(Para.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, HR, HR.L, HR.U)

save(out.carbon.tot.ses, file = here('output','analysis','out-carbon-tot-ses.RDa'))
```

```{r Output Carbon TOR SES-Adj}
out.carbon.tor.ses <- read.csv(here('output','analysis','tot_ses.csv')) %>% 
  mutate(Parameter = gsub(".*Yr_", "", Predictor),
         Parameter = as.factor(Parameter)) %>% 
  merge((carbon.method.1yr.iqr %>% filter(Method == "TOR"))) %>% 
  merge(Para.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, HR, HR.L, HR.U)

save(out.carbon.tor.ses, file = here('output','analysis','out-carbon-tor-ses.RDa'))
```

```{r Output PMC Residuals}
PM.IQR <- as.double(pmc.1yr.iqr[pmc.1yr.iqr$Parameter == 'PM', 'IQR'])
load(here('output','analysis','out-pmc-ses.RDa'))
PM <- out.pmc.ses %>% 
  filter(Index == "PM2.5")

out.pmc.resid <- read.csv(here('output','analysis','pmc_resid.csv')) %>% 
  mutate(Parameter = gsub(".*PM_", "", Predictor),
         Parameter = as.factor(Parameter)) %>% 
  merge(Para.to.Index) %>% 
  mutate(Index = paste0('-', Index)) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * PM.IQR),3),
         HR.L = round(exp((Coef-1.96*std) * PM.IQR),3),
         HR.U = round(exp((Coef+1.96*std) * PM.IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, HR, HR.L, HR.U) %>% 
  add_row(PM)

save(out.pmc.resid, file = here('output','analysis','out-pmc-resid.RDa'))
```

```{r Output Factors Residuals}
PM.IQR <- as.double(pmc.1yr.iqr[pmc.1yr.iqr$Parameter == 'PM', 'IQR'])
load(here('output','analysis','out-pmc-ses.RDa'))
PM <- out.pmc.ses %>% 
  filter(Index == "PM2.5")

out.fac.resid <- read.csv(here('output','analysis','fac_resid.csv')) %>% 
  mutate(Parameter = gsub(".*PM_", "", Predictor),
         Parameter = as.factor(Parameter)) %>% 
  merge(Factor.ex.ec.to.Index) %>% 
  mutate(Index = paste0('-', Index)) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * PM.IQR),3),
         HR.L = round(exp((Coef-1.96*std) * PM.IQR),3),
         HR.U = round(exp((Coef+1.96*std) * PM.IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, HR, HR.L, HR.U) %>% 
  add_row(PM)

save(out.fac.resid, file = here('output','analysis','out-fac-resid.RDa'))
```

```{r Output PMC Race}
out.pmc.race = data.frame()
for (i in c("A", "B", "H", "W")){
  dt <- read.csv(here('output','analysis',paste('pmc_race',i,'.csv', sep= ""))) %>% 
    mutate(Race = i)
  out.pmc.race  <-  rbind(out.pmc.race, dt)
}

out.pmc.race <- out.pmc.race %>% 
  mutate(Parameter = gsub(".*Yr_", "", Predictor),
         Parameter = as.factor(Parameter)) %>% 
  merge(pmc.1yr.iqr) %>% 
  merge(Para.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, Race, HR, HR.L, HR.U)

save(out.pmc.race, file = here('output','analysis','out-pmc-race.RDa'))
```

```{r Output PMC Urbanicity}
out.pmc.urbanicity = data.frame()
for (i in c("urban", "rural")){
  dt <- read.csv(here('output','analysis',paste('pmc_',i,'.csv', sep= ""))) %>% 
    mutate(Loc.Set = i)
  out.pmc.urbanicity  <-  rbind(out.pmc.urbanicity, dt)
}

out.pmc.urbanicity <- out.pmc.urbanicity %>% 
  mutate(Parameter = gsub(".*Yr_", "", Predictor),
         Parameter = as.factor(Parameter)) %>% 
  merge(pmc.1yr.iqr) %>% 
  merge(Para.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, Loc.Set, HR, HR.L, HR.U)

save(out.pmc.urbanicity, file = here('output','analysis','out-pmc-urbanicity.RDa'))
```

```{r Output PMC Urban - SES}
out.pmc.sescat = data.frame()
for (i in c("L", "M", "H")){
  dt <- read.csv(here('output','analysis',paste('pmc_urban_ses',i,'.csv', sep= ""))) %>% 
    mutate(SES.Cat = i)
  out.pmc.sescat  <-  rbind(out.pmc.sescat, dt)
}

out.pmc.sescat <- out.pmc.sescat %>% 
  mutate(Parameter = gsub(".*Yr_", "", Predictor),
         Parameter = as.factor(Parameter)) %>% 
  merge(pmc.1yr.iqr) %>% 
  merge(Para.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, SES.Cat, HR, HR.L, HR.U)

save(out.pmc.sescat, file = here('output','analysis','out-pmc-sescat.RDa'))
```

```{r Output PMC Region}
out.pmc.region = data.frame()
for (i in c("west", "midwest", "south", "northeast")){
  dt <- read.csv(here('output','analysis',paste('pmc_',i,'.csv', sep= ""))) %>% 
    mutate(Region = i)
  out.pmc.region  <-  rbind(out.pmc.region, dt)
}

out.pmc.region <- out.pmc.region %>% 
  mutate(Parameter = gsub(".*Yr_", "", Predictor),
         Parameter = as.factor(Parameter)) %>% 
  merge(pmc.1yr.iqr) %>% 
  merge(Para.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, Region, HR, HR.L, HR.U)

save(out.pmc.region, file = here('output','analysis','out-pmc-region.RDa'))
```

```{r Output PMC Age}
out.pmc.age = data.frame()
for (i in c("LE75", "M75")){
  dt <- read.csv(here('output','analysis',paste('pmc_age',i,'.csv', sep= ""))) %>% 
    mutate(Age = i)
  out.pmc.age  <-  rbind(out.pmc.age, dt)
}

out.pmc.age <- out.pmc.age %>% 
  mutate(Parameter = gsub(".*Yr_", "", Predictor),
         Parameter = as.factor(Parameter)) %>% 
  merge(pmc.1yr.iqr) %>% 
  merge(Para.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, Age, HR, HR.L, HR.U)

save(out.pmc.age, file = here('output','analysis','out-pmc-age.RDa'))
```

```{r Output PMC Sex}
out.pmc.sex = data.frame()
for (i in c("F", "M")){
  dt <- read.csv(here('output','analysis',paste('pmc_sex',i,'.csv', sep= ""))) %>% 
    mutate(Sex = i)
  out.pmc.sex  <-  rbind(out.pmc.sex, dt)
}

out.pmc.sex <- out.pmc.sex %>% 
  mutate(Parameter = gsub(".*Yr_", "", Predictor),
         Parameter = as.factor(Parameter)) %>% 
  merge(pmc.1yr.iqr) %>% 
  merge(Para.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, Sex, HR, HR.L, HR.U)

save(out.pmc.sex, file = here('output','analysis','out-pmc-sex.RDa'))
```

```{r Output PMC BRFSS}
out.pmc.brfss = data.frame()
for (i in c("brfss0", "brfss")){
  dt <- read.csv(here('output','analysis',paste('pmc_',i,'.csv', sep= ""))) %>% 
    mutate(BRFSS = i)
  out.pmc.brfss  <-  rbind(out.pmc.brfss, dt)
}

out.pmc.brfss <- out.pmc.brfss %>% 
  mutate(Parameter = gsub(".*Yr_", "", Predictor),
         Parameter = as.factor(Parameter)) %>% 
  merge(pmc.1yr.iqr) %>% 
  merge(Para.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, BRFSS, HR, HR.L, HR.U)

save(out.pmc.brfss, file = here('output','analysis','out-pmc-brfss.RDa'))
```

```{r Output PMC bufferzone}
out.pmc.buffer = data.frame()
for (i in c("6", "12", "24")){
  dt <- read.csv(here('output','analysis',paste('pmc_bz',i,'.csv', sep= ""))) %>% 
    mutate(Buffer = i)
  out.pmc.buffer  <-  rbind(out.pmc.buffer, dt)
}

out.pmc.buffer <- out.pmc.buffer %>% 
  mutate(Parameter = gsub(".*Yr_", "", Predictor),
         Parameter = as.factor(Parameter)) %>% 
  merge(pmc.1yr.iqr) %>% 
  merge(Para.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, Buffer, HR, HR.L, HR.U)

save(out.pmc.buffer, file = here('output','analysis','out-pmc-buffer.RDa'))
```

```{r Output PMC Exposure window}
out.pmc.expo = data.frame()
for (i in c("1", "2", "3", "4", "5")){
  dt <- read.csv(here('output','analysis',paste('pmc_',i,'yr','.csv', sep= ""))) %>% 
    mutate(Window = i)
  out.pmc.expo  <-  rbind(out.pmc.expo, dt)
}

out.pmc.expo <- out.pmc.expo %>% 
  mutate(Parameter = gsub(".*Yr_", "", Predictor),
         Parameter = as.factor(Parameter)) %>% 
  merge(pmc.1yr.iqr) %>% 
  merge(Para.to.Index) %>% 
  merge(Outcome.to.Cause) %>% 
  mutate(HR = round(exp(Coef * IQR),3),
         HR.L = round(exp((Coef-1.96*std) * IQR),3),
         HR.U = round(exp((Coef+1.96*std) * IQR),3)) %>% 
  arrange(Config) %>% 
  select(Index, Cause, Window, HR, HR.L, HR.U)

save(out.pmc.expo, file = here('output','analysis','out-pmc-expo.RDa'))
```