### 2021-02-14 code/data-viz/forest-plots-color.Rmd

---
title: "PM Component Paper Plots"
author: "Fatemeh Kazemi"
date: "01-19-2021"
---

### plots subgroups using different colors
###color codes: https://www.schemecolor.com/british-airways-website.php

```{r}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r Load packages}
  library(tidyverse)
  library(here)
  library(ggthemes)
```

```{r Call Functions}
source(here('code','data-viz','forest-plot-functions.R'))
```

```{r Load data}
load(here('output','analysis','out-pmc-base.RDa'))
load(here('output','analysis','out-pmc-ses.RDa'))
load(here('output','analysis','out-fac-base.RDa'))
load(here('output','analysis','out-fac-ses.RDa'))
load(here('output','analysis','out-fac-base-ec.RDa'))
load(here('output','analysis','out-fac-ses-ec.RDa'))
load(here('output','analysis','out-pmc-resid.RDa'))
load(here('output','analysis','out-fac-resid.RDa'))
load(here('output','analysis','out-pmc-age.RDa'))
load(here('output','analysis','out-fac-age.RDa'))
load(here('output','analysis','out-pmc-sex.RDa'))
load(here('output','analysis','out-fac-sex.RDa'))
load(here('output','analysis','out-pmc-race.RDa'))
load(here('output','analysis','out-fac-race.RDa'))
load(here('output','analysis','out-pmc-urbanicity.RDa'))
load(here('output','analysis','out-fac-urbanicity.RDa'))
load(here('output','analysis','out-pmc-sescat.RDa'))
load(here('output','analysis','out-pmc-region.RDa'))
load(here('output','analysis','out-pmc-resid-region.RDa'))
load(here('output','analysis','out-fac-region.RDa'))
load(here('output','analysis','out-pmc-brfss.RDa'))
load(here('output','analysis','out-pmc-buffer.RDa'))
load(here('output','analysis','out-pmc-expo.RDa'))
```

```{r set theme 1 - single group plots}
ggtheme.single <- theme_bw() + 
  theme(panel.spacing.x = unit(0.25, 'lines'),
        panel.border = element_rect(fill = NA, 'lines', color = "gray25"),
        strip.text.x = element_text(size = 10, face = 'bold', color = "black"),
        strip.background = element_rect(fill = "#B9CFED"),
        axis.text.x = element_text(size = 8, color = 'black'),
        axis.text.y = element_text(size = 8, face = 'bold', color = 'black'),
  )
```

```{r set theme 2 - components - effect modification, subgroups, multiple groups plots}
ggtheme.multi.1 <- theme_classic() + 
  theme(panel.spacing.x = unit(0.25,'lines'),
        panel.spacing.y = unit(c(0.25,0,0.25,0,0.25,0,0.25,0,0.25,0),'lines'),
        panel.border = element_rect(fill = NA, 'lines', color = "gray25"),
        strip.text.x = element_text(size = 10, face = 'bold', color = "black"),
        strip.text.y = element_text(size = 10, face = "bold", color = "black", angle = 180),
        strip.text.y.left = element_text(angle = 0),
        strip.background = element_rect(fill = "#B9CFED"),
        axis.text.x = element_text(size = 8, color = 'black'),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = 'bottom',
        legend.title = element_blank(),
        legend.text = element_text(size = 10, face = "bold"),
        legend.key.width = unit(0.25, units = 'in'))
```

```{r set theme 3 - factors - effect modification, subgroups, multiple groups plots}
ggtheme.multi.2 <- theme_classic() + 
  theme(panel.spacing.x = unit(0.25,'lines'),
        panel.spacing.y = unit(c(0.25),'lines'),
        panel.border = element_rect(fill = NA, 'lines', color = "gray25"),
        strip.text.x = element_text(size = 10, face = 'bold', color = "black"),
        strip.text.y = element_text(size = 10, face = "bold", color = "black", angle = 180),
        strip.text.y.left = element_text(angle = 0),
        strip.background = element_rect(fill = "#B9CFED"),
        axis.text.x = element_text(size = 8, color = 'black'),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = 'bottom',
        legend.title = element_blank(),
        legend.text = element_text(size = 10, face = "bold"),
        legend.key.width = unit(0.25, units = 'in'))
```

```{r}
cuz <- c("All-Cause", "Cardiovascular", "IHD", "CBV", "CHF",
         "Respiratory", "COPD", "Pneumonia", 
         "Cancer", "Lung Cancer")

cuz.top.row <-  c ("All-Cause", "Cardiovascular", "IHD", "CBV", "CHF")
cuz.bot.row <- c ("Respiratory", "COPD", "Pneumonia", "Cancer", "Lung Cancer")

cmp.main <- c("Mn", "Fe", "V", "Ni", "Cl", "Na", "SO4", "Se", 
              "Cu", "K", "OC", "NO3", "Zn", "Pb", "Ca", "Si", "PM2.5")
cmp.multi <- c("PM2.5","Si", "Ca", "Pb", "Zn", "NO3", "OC", "Se", "SO4", 
               "Ni", "V")

So.main <- c("Steel", "Oil", "Salt", "Coal", 
         "Biomass", "Traffic", "Metals", "Soil")
So.multi <- c("Soil", "Metals", "Traffic",
              "Coal", "Oil")

So.ec <- c("Steel", "Coal", "Oil", "Salt",
         "Biomass", "Metals", "Soil", "Traffic")

cmp.res <- c("-V", "-Ni","-SO4", "-Se", "-OC", "-NO3", 
         "-Zn", "-Pb", "-Ca", "-Si", "PM2.5")
So.res <- c("-Oil", "-Coal", 
         "-Traffic", "-Metals", "-Soil", "PM2.5")

cmp.multi.res <- c("PM2.5","-Si", "-Ca", "-Pb", "-Zn", "-NO3", "-OC",
                   "-Se", "-SO4", "-Ni", "-V")
```

```{r - Components Base}
data <- out.pmc.base %>% 
  filter(Index %in% cmp.main,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = cmp.main),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.cols <- c("#EB2226")

Plot.Single(dt = data, ggtheme = ggtheme.single) +
ggsave(here('output','plot','pmc-base.png'), width = 6.5, height = 6.5)
```

```{r - Components SES-Adj}
data <- out.pmc.ses %>% 
  filter(Index %in% cmp.main,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = cmp.main),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.cols <- c("#EB2226")

Plot.Single(dt = data, ggtheme = ggtheme.single) +
ggsave(here('output','plot','pmc-ses.png'), width = 6.5, height = 6.5)
```

```{r - Factor Base Ex. EC}
data <- out.fac.base %>% 
  filter(Index %in% So.main,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = So.main),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.cols <- c("#EB2226")

Plot.Single(dt = data, ggtheme = ggtheme.single) +
ggsave(here('output','plot','fac-base.png'), width = 6.5, height = 6.5)
```

```{r - Factor SES-Adj Ex. EC}
data <- out.fac.ses %>% 
  filter(Index %in% So.main,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = So.main),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.cols <- c("#EB2226")

Plot.Single(dt = data, ggtheme = ggtheme.single) +
ggsave(here('output','plot','fac-ses.png'), width = 6.5, height = 6.5)
```

```{r - Factor Inc. EC Base}
data <- out.fac.base.ec %>% 
  filter(Index %in% So.ec,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = So.ec),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.cols <- c("#EB2226")

Plot.Single(dt = data, ggtheme = ggtheme.single) +
ggsave(here('output','plot','fac-base-ec.png'), width = 6.5, height = 6.5)
```

```{r - Factor Inc. EC SES-Adj}
data <- out.fac.ses.ec %>% 
  filter(Index %in% So.ec,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = So.ec),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.cols <- c("#EB2226")

Plot.Single(dt = data, ggtheme = ggtheme.single) +
ggsave(here('output','plot','fac-ses-ec.png'), width = 6.5, height = 6.5)
```

```{r - Components - PM Residual}
data <- out.pmc.resid %>% 
  filter(Index %in% cmp.res,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = cmp.res),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.cols <- c("#EB2226")

Plot.Single(dt = data, ggtheme = ggtheme.single) +
ggsave(here('output','plot','pmc-resid.png'), width = 6.5, height = 6.5)
```

```{r - Factor - PM Residual}
data <- out.fac.resid %>% 
  filter(Index %in% So.res,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = So.res),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.cols <- c("#EB2226")

Plot.Single(dt = data, ggtheme = ggtheme.single) +
ggsave(here('output', 'plot','fac-resid.png'), width = 6.5, height = 6.5)
```

```{r - Components Age}
data <- out.pmc.age %>% 
  filter(Index %in% cmp.multi,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = cmp.multi),
         Tag = ifelse(Age == "LE75", "b", "a"),
         Tag = as.factor(Tag),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.labs <- c("> 75 years", "<= 75 years")
plot.cols <- c("#EB2226", "#075AAA")

Plot.Multi.Color(dt = data, ggtheme = ggtheme.multi.1, n = 1) +
ggsave(here('output', 'plot','pmc-age.png'), width = 6.5, height = 7.5)
```

```{r - Factor Age}
data <- out.fac.age %>% 
  filter(Index %in% So.multi,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = So.multi),
         Tag = ifelse(Age == "LE75", "b", "a"),
         Tag = as.factor(Tag),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.labs <- c("> 75 years", "<= 75 years")
plot.cols <- c("#EB2226", "#075AAA")

Plot.Multi.Color(dt = data, ggtheme = ggtheme.multi.2, n = 1) +
ggsave(here('output', 'plot','fac-age.png'), width = 6.5, height = 6.5)
```

```{r - Components Sex}
data <- out.pmc.sex %>% 
  filter(Index %in% cmp.multi,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = cmp.multi),
         Tag = ifelse(Sex == "F", "b", "a"),
         Tag = as.factor(Tag),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.labs <- c("Male", "Female")
plot.cols <- c("#EB2226", "#075AAA")

Plot.Multi.Color(dt = data, ggtheme = ggtheme.multi.1, n = 1) +
ggsave(here('output', 'plot','pmc-sex.png'), width = 6.5, height = 7.5)
```

```{r - Factor Sex}
data <- out.fac.sex %>% 
  filter(Index %in% So.multi,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = So.multi),
         Tag = ifelse(Sex == "F", "b", "a"),
         Tag = as.factor(Tag),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.labs <- c("Male", "Female")
plot.cols <- c("#EB2226", "#075AAA")

Plot.Multi.Color(dt = data, ggtheme = ggtheme.multi.2, n = 1) +
ggsave(here('output', 'plot','fac-sex.png'), width = 6.5, height = 6.5)
```

```{r - Components Race}
data <- out.pmc.race %>% 
  filter(Index %in% cmp.multi,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = cmp.multi),
         Tag = ifelse(Race == "A", "d",
                      ifelse(Race == "B", "c",
                             ifelse(Race == "H", "b", "a"))),
         Tag = as.factor(Tag),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.labs <- c("White", "Hispanic", "Black", "Asian")
plot.cols <- c("#FF8200", "#075AAA", "#EB2226", "#01742F")

Plot.Multi.Color(dt = data, ggtheme = ggtheme.multi.1, n = 1) +
ggsave(here('output', 'plot','pmc-race.png'), width = 6.5, height = 7.5)
```

```{r - Factor Race}
data <- out.fac.race %>% 
  filter(Index %in% So.multi,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = So.multi),
         Tag = ifelse(Race == "A", "d",
                      ifelse(Race == "B", "c",
                             ifelse(Race == "H", "b", "a"))),
         Tag = as.factor(Tag),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.labs <- c("White", "Hispanic", "Black", "Asian")
plot.cols <- c("#FF8200", "#075AAA", "#EB2226", "#01742F")

Plot.Multi.Color(dt = data, ggtheme = ggtheme.multi.2, n = 1) +
ggsave(here('output', 'plot','fac-race.png'), width = 6.5, height = 6.5)
```

```{r Components Urban}
data <- out.pmc.urbanicity %>% 
  filter(Index %in% cmp.multi,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = cmp.multi),
         Tag = ifelse(Loc.Set == "urban", "b", "a"),
         Tag = as.factor(Tag),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.labs <- c("Rural", "Urban")
plot.cols <- c("#EB2226", "#075AAA")

Plot.Multi.Color(dt = data, ggtheme = ggtheme.multi.1, n = 1) +
ggsave(here('output', 'plot','pmc-urbanicity.png'), width = 6.5, height = 7.5)
```

```{r - Factor Urbaniity}
data <- out.fac.urbanicity %>% 
  filter(Index %in% So.multi,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = So.multi),
         Tag = ifelse(Loc.Set == "urban", "b", "a"),
         Tag = as.factor(Tag),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.labs <- c("Rural", "Urban")
plot.cols <- c("#EB2226", "#075AAA")

Plot.Multi.Color(dt = data, ggtheme = ggtheme.multi.2, n = 1) +
ggsave(here('output', 'plot','fac-urbanicity.png'), width = 6.5, height = 6.5)
```

```{r Components SES-Urban}
data <- out.pmc.sescat %>% 
  filter(Index %in% cmp.multi,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = cmp.multi),
         Tag = ifelse(SES.Cat == "L", "c",
                      ifelse(SES.Cat == "M", "b", "a")),
         Tag = as.factor(Tag),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.labs <- c("High", "Middle", "Low")
plot.cols <-  c("#01742F", "#075AAA", "#EB2226")

Plot.Multi.Color(dt = data, ggtheme = ggtheme.multi.1, n = 1) +
ggsave(here('output', 'plot','pmc-sescat.png'), width = 6.5, height = 7.5)
```

```{r - Components Region}
data <- out.pmc.region %>% 
  filter(Index %in% cmp.multi,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = cmp.multi),
         Tag = ifelse(Region == "west", "d",
                      ifelse(Region == "midwest", "c",
                             ifelse(Region == "south", "b", "a"))),
         Tag = as.factor(Tag),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.labs <- c("Northeast", "South", "Midwest", "West")
plot.cols <- c("#07f900", "#0433ff", "#ffa501", "#a034f0")

Plot.Multi.Color(dt = data, ggtheme = ggtheme.multi.1, n = 1) +
ggsave(here('output', 'plot','pmc-region.png'), width = 6.5, height = 7.5)
```

```{r - Factor Region}
data <- out.fac.region %>% 
  filter(Index %in% So.multi,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = So.multi),
  Tag = ifelse(Region == "west", "d",
               ifelse(Region == "midwest", "c",
                      ifelse(Region == "south", "b", "a"))),
         Tag = as.factor(Tag),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.labs <- c("Northeast", "South", "Midwest", "West")
plot.cols <- c("#07f900", "#0433ff", "#ffa501", "#a034f0")

Plot.Multi.Color(dt = data, ggtheme = ggtheme.multi.2, n = 1) +
ggsave(here('output', 'plot','fac-region.png'), width = 6.5, height = 6.5)
```

```{r - Components Residual Region}
data <- out.pmc.resid.region %>% 
  filter(Index %in% cmp.multi.res,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = cmp.multi.res),
         Tag = ifelse(Region == "west", "d",
                      ifelse(Region == "midwest", "c",
                             ifelse(Region == "south", "b", "a"))),
         Tag = as.factor(Tag),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.labs <- c("Northeast", "South", "Midwest", "West")
plot.cols <- c("#07f900", "#0433ff", "#ffa501", "#a034f0")

Plot.Multi.Color(dt = data, ggtheme = ggtheme.multi.1, n = 1) +
ggsave(here('output', 'plot','pmc-resid-region.png'),width = 6.5, height = 7.5)
```

```{r - Components BRFSS}
data <- out.pmc.brfss %>% 
  filter(Index %in% cmp.multi,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = cmp.multi),
         Tag = ifelse(BRFSS == "brfss0", "b", "a"),
         Tag = as.factor(Tag),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.labs <- c("BRFSS-Adjusted", "Base")
plot.cols <- c("#EB2226", "#075AAA")

Plot.Multi.Color(dt = data, ggtheme = ggtheme.multi.1, n = 1) +
ggsave(here('output', 'plot','pmc-brfss.png'), width = 6.5, height = 7.5)
```

```{r Components Buffer}
data <- out.pmc.buffer %>% 
  filter(Index %in% cmp.multi,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = cmp.multi),
         Tag = ifelse(Buffer == 6, "c",
                      ifelse(Buffer == 12, "b", "a")),
         Tag = as.factor(Tag),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.labs <- c("24-miles", "12-miles", "6-miles") 
plot.cols <- c("#EB2226", "#075AAA", "#01742F")

Plot.Multi.Color(dt = data, ggtheme = ggtheme.multi.1, n = 1) +
ggsave(here('output', 'plot','pmc-buffer.png'), width = 6.5, height = 7.5)
```

```{r Components Exposure window}
data <- out.pmc.expo %>% 
  filter(Index %in% cmp.multi,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz),
         Index = factor(Index, levels = cmp.multi),
         Tag = ifelse(Window == 1, "e",
                      ifelse(Window == 2, "d",
                             ifelse(Window == 3, "c",
                                    ifelse(Window == 4, "b", "a")))),
         Tag = as.factor(Tag),
         sig = ifelse(HR.U < 1 , 16, #16-prtc
                      ifelse((HR.L<=1 & HR.U>=1), 1, 16)) #1-null # 16-sigpos
  )

plot.labs <- c("5-yr", "4-yr","3-yr","2-yr","1-yr")
plot.cols <- c("#EB2226","#FF8200","#9D178D","#075AAA","#01742F")

Plot.Multi.Color(dt = data, ggtheme = ggtheme.multi.1, n = 1) + 
ggsave(here('output', 'plot','pmc-expo.png'), width = 6.5, height = 8.5)
```