### 2021-05-31 code/data-viz/manuscript-tables.Rmd

---
title: "Tables for PM2.5 Paper Manuscript"
author: "Fatemeh Kazemi"
date: "3/28/2021"
output:
  html_document:
    df_print: paged
---

```{r}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r load packages}
library(tidyverse)
library(here)
library(knitr)
library(kableExtra)
```

```{r Load data}
load(here('output','analysis','out-pmc-base.RDa'))
load(here('output','analysis','out-pmc-ses.RDa'))
load(here('output','analysis','out-fac-base.RDa'))
load(here('output','analysis','out-fac-ses.RDa'))
#load(here('output','analysis','out-fac-base-ec.RDa'))
#load(here('output','analysis','out-fac-ses-ec.RDa'))
load(here('output','analysis','out-pmc-resid.RDa'))
load(here('output','analysis','out-fac-resid.RDa'))
load(here('output','analysis','out-pmc-age.RDa'))
load(here('output','analysis','out-fac-age.RDa'))
load(here('output','analysis','out-pmc-sex.RDa'))
load(here('output','analysis','out-fac-sex.RDa'))
load(here('output','analysis','out-pmc-race.RDa'))
load(here('output','analysis','out-fac-race.RDa'))
load(here('output','analysis','out-pmc-urbanicity.RDa'))
load(here('output','analysis','out-fac-urbanicity.RDa'))
load(here('output','analysis','out-pmc-sescat.RDa'))
load(here('output','analysis','out-pmc-region.RDa'))
load(here('output','analysis','out-pmc-resid-region.RDa'))
load(here('output','analysis','out-fac-region.RDa'))
load(here('output','analysis','out-pmc-brfss.RDa'))
load(here('output','analysis','out-pmc-buffer.RDa'))
load(here('output','analysis','out-pmc-expo.RDa'))
```

```{r}
cuz <- c("All-Cause", "Cardiovascular", "IHD", "CBV", "CHF",
         "Respiratory", "COPD", "Pneumonia", 
         "Cancer", "Lung Cancer")

cuz.top.row <-  c ("All-Cause", "Cardiovascular", "IHD", "CBV", "CHF")
cuz.bot.row <- c ("Respiratory", "COPD", "Pneumonia", "Cancer", "Lung Cancer")

cmp.main <- c("PM2.5", "Si","Ca", "Pb","Zn", "NO3","OC", "K","Cu",
              "Se","SO4", "Na","Cl", "Ni","V", "Fe","Mn")

cmp.multi <- c("PM2.5","Si", "Ca", "Pb", "Zn", "NO3", "OC", "Se", "SO4", 
               "Ni", "V") 
cmp.m2 <- c("PM2.5", "PM2.5", "Si", "Si", "Ca", "Ca", "Pb", "Pb", "Zn", "Zn",
            "NO3", "NO3", "OC", "OC", "Se", "Se", "SO4", "SO4", "Ni", "Ni", "V", "V")

So.main <- c("Soil", "Metals", "Traffic", "Biomass",
             "Coal", "Salt", "Oil", "Steel")

So.multi <- c("Soil", "Metals", "Traffic", "Coal", "Oil")

cmp.res <- c("PM2.5", "-Si","-Ca", "-Pb","-Zn", "-NO3","-OC",
             "-Se","-SO4", "-Ni","-V")

So.res <- c("PM2.5", "-Soil", "-Metals", "-Traffic", "-Coal", "-Oil")

cmp.multi.res <- c("PM2.5","-Si", "-Ca", "-Pb", "-Zn", "-NO3", "-OC",
                   "-Se", "-SO4", "-Ni", "-V")

Race <- c("A", "B", "H", "W")
AgeC <- c("LE75", "M75")
```

# Figure 1a
```{r}
table_fk <- function(data, cmp, cuz){
  opts <- options(knitr.kable.NA = "0")
  kbl(data %>% 
  filter(Index %in% cmp,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz)) %>% 
  mutate_at(.vars = c("HR", "HR.L", "HR.U"), ~sprintf('%.3f', .x)) %>% 
  mutate(HR.Sum = paste(HR," (", HR.L,"-", HR.U,")", sep = "")) %>% 
  select(-(HR:HR.U)) %>% 
  pivot_wider(names_from = "Cause",
              values_from = "HR.Sum") %>% 
  slice(match(cmp, Index)),
  align = "c", booktabs = TRUE) %>%
    kable_classic(full_width = F, html_font = "Times New Roman", font_size = 8) %>%
    group_rows("Soil", 2, 3) %>%   # same result with group 1=
    group_rows("Metals", 4, 5) %>% 
    group_rows("Traffic", 6, 7) %>% 
    group_rows("Biomass", 8, 9) %>% 
    group_rows("Coal", 10, 11) %>%
    group_rows("Salt", 12, 13) %>% 
    group_rows("Oil", 14, 15) %>% 
    group_rows("Steel", 16, 17) %>% 
    kable_styling() %>% 
    row_spec(0, font_size = 8)
}

table_fk(data = out.pmc.ses, cmp = cmp.main, cuz = cuz.top.row)
```

# Figure 1a
```{r}
table_fk(data = out.pmc.ses, cmp = cmp.main, cuz = cuz.bot.row)
```

# Figure 1b
```{r}
table_fk <- function(data, cmp, cuz){
  opts <- options(knitr.kable.NA = "0")
  kbl(data %>% 
  filter(Index %in% cmp,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz)) %>% 
  mutate_at(.vars = c("HR", "HR.L", "HR.U"), ~sprintf('%.3f', .x)) %>% 
  mutate(HR.Sum = paste(HR," (", HR.L,"-", HR.U,")", sep = "")) %>% 
  select(-(HR:HR.U)) %>% 
  pivot_wider(names_from = "Cause",
              values_from = "HR.Sum") %>% 
  slice(match(cmp, Index)),
  align = "c", booktabs = TRUE) %>%
    kable_classic(full_width = F, html_font = "Times New Roman", font_size = 11) %>%
    kable_styling() %>% 
    row_spec(0, font_size = 9)
}

table_fk(data = out.fac.ses, cmp = So.main, cuz = cuz.top.row)
```

# Figure 1b
```{r}
table_fk(data = out.fac.ses, cmp = So.main, cuz = cuz.bot.row)
```

# Figure 2a
```{r}
table_fk <- function(data, cmp, cuz){
  opts <- options(knitr.kable.NA = "0")
  kbl(data %>% 
  filter(Index %in% cmp,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz)) %>% 
  mutate_at(.vars = c("HR", "HR.L", "HR.U"), ~sprintf('%.3f', .x)) %>% 
  mutate(HR.Sum = paste(HR," (", HR.L,"-", HR.U,")", sep = "")) %>% 
  select(-(HR:HR.U)) %>% 
  pivot_wider(names_from = "Cause",
              values_from = "HR.Sum") %>% 
  slice(match(cmp, Index)),
  align = "c", booktabs = TRUE) %>%
    kable_classic(full_width = F, html_font = "Times New Roman", font_size = 11) %>%
    group_rows("Soil", 2, 3) %>%   # same result with group 1=
    group_rows("Metals", 4, 5) %>% 
    group_rows("Traffic", 6, 7) %>% 
    group_rows("Coal", 8, 9) %>%
    group_rows("Oil", 10, 11) %>% 
    kable_styling() %>% 
    row_spec(0, font_size = 9)
}

table_fk(data = out.pmc.resid, cmp = cmp.res, cuz = cuz.top.row)
```

# Figure 2a
```{r}
table_fk(data = out.pmc.resid, cmp = cmp.res, cuz = cuz.bot.row)
```

# Figure 2b
```{r}
table_fk <- function(data, cmp, cuz){
  opts <- options(knitr.kable.NA = "0")
  kbl(data %>% 
  filter(Index %in% cmp,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz)) %>% 
  mutate_at(.vars = c("HR", "HR.L", "HR.U"), ~sprintf('%.3f', .x)) %>% 
  mutate(HR.Sum = paste(HR," (", HR.L,"-", HR.U,")", sep = "")) %>% 
  select(-(HR:HR.U)) %>% 
  pivot_wider(names_from = "Cause",
              values_from = "HR.Sum") %>% 
  slice(match(cmp, Index)),
  align = "c", booktabs = TRUE) %>%
    kable_classic(full_width = F, html_font = "Times New Roman", font_size = 11) %>%
    kable_styling() %>% 
    row_spec(0, font_size = 9)
}

table_fk(data = out.fac.resid, cmp = So.res, cuz = cuz.top.row)
```

# Figure 2b
```{r}
table_fk(data = out.fac.resid, cmp = So.res, cuz = cuz.bot.row)
```

# Figure x
```{r}
table_fk <- function(data, cmp, cuz, em){
  opts <- options(knitr.kable.NA = "0")
  kbl(data %>% 
  filter(!Index %in% c("Na", "Cl")) %>% 
  filter(Index %in% cmp,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz)) %>% 
  mutate_at(.vars = c("HR", "HR.L", "HR.U"), ~sprintf('%.3f', .x)) %>% 
  mutate(HR.Sum = paste(HR," (", HR.L,"-", HR.U,")", sep = "")) %>% 
  select(-(HR:HR.U)) %>% 
  pivot_wider(names_from = "Cause",
              values_from = "HR.Sum") %>% 
    arrange(factor(Index, levels = cmp), Age),
  align = "c", booktabs = TRUE) %>%
    kable_classic(full_width = F, html_font = "Times New Roman", font_size = 8) %>%
    group_rows("Soil", 3, 6) %>%   # same result with group 1=
    group_rows("Metals", 7, 10) %>% 
    group_rows("Traffic", 11, 14) %>% 
    group_rows("Coal", 15, 18) %>% 
    group_rows("Oil", 19, 22) %>%
    kable_styling() %>% 
    row_spec(0, font_size = 8)
}

table_fk(data = out.pmc.age, cmp = cmp.multi, cuz = cuz.top.row)
```

```{r}
table_fk(data = out.pmc.age, cmp = cmp.multi, cuz = cuz.bot.row)
```

# Figure x2
```{r}
table_fk <- function(data, cmp, cuz){
  opts <- options(knitr.kable.NA = "0")
  kbl(data %>% 
  filter(Index %in% cmp,
         Cause %in% cuz) %>% 
  mutate(Cause = factor(Cause, levels = cuz)) %>% 
  mutate_at(.vars = c("HR", "HR.L", "HR.U"), ~sprintf('%.3f', .x)) %>% 
  mutate(HR.Sum = paste(HR," (", HR.L,"-", HR.U,")", sep = "")) %>% 
  select(-(HR:HR.U)) %>% 
  pivot_wider(names_from = "Cause",
              values_from = "HR.Sum") %>% 
    arrange(factor(Index, levels = cmp), Age),
  align = "c", booktabs = TRUE) %>%
    kable_classic(full_width = F, html_font = "Times New Roman", font_size = 11) %>%
    group_rows("Soil", 3, 6) %>%   # same result with group 1=
    group_rows("Metals", 7, 10) %>% 
    group_rows("Traffic", 11, 14) %>% 
    group_rows("Coal", 15, 18) %>% 
    group_rows("Oil", 19, 22) %>%
    kable_styling() %>% 
    row_spec(0, font_size = 9)
}

table_fk(data = out.fac.age, cmp = So.multi, cuz = cuz.top.row)
```

```{r}
table_fk(data = out.fac.age, cmp = So.multi, cuz = cuz.bot.row)
```