### 2021-06-03 code/data-viz/manuscript-tables.Rmd

---
title: "Tables for PM2.5 Components Paper Manuscript"
author: "Fatemeh Kazemi"
date: "05/31/2021"
output:
  html_document:
    df_print: paged
---

```{r}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r load packages}
library(tidyverse)
library(here)
library(flextable)
library(officer)
```

```{r Call Functions}
source(here('code','data-viz','manuscript-table-functions.R'))
```

```{r Load data}
load(here('output','analysis','out-pmc-ses.RDa'))
load(here('output','analysis','out-fac-ses.RDa'))
load(here('output','analysis','out-pmc-resid.RDa'))
load(here('output','analysis','out-fac-resid.RDa'))
load(here('output','analysis','out-pmc-age.RDa'))
load(here('output','analysis','out-fac-age.RDa'))
load(here('output','analysis','out-pmc-sex.RDa'))
load(here('output','analysis','out-fac-sex.RDa'))
load(here('output','analysis','out-pmc-race.RDa'))
load(here('output','analysis','out-fac-race.RDa'))
load(here('output','analysis','out-pmc-urbanicity.RDa'))
load(here('output','analysis','out-fac-urbanicity.RDa'))
load(here('output','analysis','out-pmc-sescat.RDa'))
load(here('output','analysis','out-pmc-region.RDa'))
load(here('output','analysis','out-pmc-resid-region.RDa'))
load(here('output','analysis','out-fac-region.RDa'))
load(here('output','analysis','out-pmc-brfss.RDa'))
load(here('output','analysis','out-pmc-buffer.RDa'))
load(here('output','analysis','out-pmc-expo.RDa'))
```

```{r}
cuz <- c("All-Cause", "Cardiovascular", "IHD", "CBV", "CHF",
         "Respiratory", "COPD", "Pneumonia", 
         "Cancer", "Lung Cancer")

cuz.top.row <-  c ("All-Cause", "Cardiovascular", "IHD", "CBV", "CHF")
cuz.bot.row <- c ("Respiratory", "COPD", "Pneumonia", "Cancer", "Lung Cancer")

cmp.main <- c("PM2.5", "Si","Ca", "Pb","Zn", "NO3","OC", "K","Cu",
              "Se","SO4", "Na","Cl", "Ni","V", "Fe","Mn")

cmp.multi <- c("PM2.5","Si", "Ca", "Pb", "Zn", "NO3", "OC", "Se", "SO4", 
               "Ni", "V") 
cmp.multi.labels <- c(expression("PM"[2.5]),"Si", "Ca", "Pb", "Zn", 
                      expression("NO"[3]^{"-"}), "OC", "Se", expression("SO"[4]^{"2-"}),
                      "Ni", "V")

So.main <- c("Soil", "Metals", "Traffic", "Biomass",
             "Coal", "Salt", "Oil", "Steel")

So.multi <- c("Soil", "Metals", "Traffic", "Coal", "Oil")

cmp.res <- c("PM2.5", "-Si","-Ca", "-Pb","-Zn", "-NO3","-OC",
             "-Se","-SO4", "-Ni","-V")

So.res <- c("PM2.5", "-Soil", "-Metals", "-Traffic", "-Coal", "-Oil")

cmp.multi.res <- c("PM2.5","-Si", "-Ca", "-Pb", "-Zn", "-NO3", "-OC",
                   "-Se", "-SO4", "-Ni", "-V")
```

```{r}
doc <- read_docx()
doc <- Table.Single(dt = out.pmc.ses, cmp = cmp.main, cuz = cuz.top.row,
                 cpt = "Figure 1", ftnt = "Continued ...")
doc <- Table.Single(dt = out.pmc.ses, cmp = cmp.main, cuz = cuz.bot.row,
                 cpt = "Figure 1, Continued", ftnt = "")
```

```{r}
doc <- Table.Single(dt = out.fac.ses, cmp = So.main, cuz = cuz.top.row,
                 cpt = "Figure 2", ftnt = "Continued ...")
doc <- Table.Single(dt = out.fac.ses, cmp = So.main, cuz = cuz.bot.row,
                 cpt = "Figure 2, Continued", ftnt = "")
```

```{r}
doc <- Table.Single(dt = out.pmc.resid, cmp = cmp.res, cuz = cuz.top.row,
                 cpt = "Figure 3", ftnt = "Continued ...")
doc <- Table.Single(dt = out.pmc.resid, cmp = cmp.res, cuz = cuz.bot.row,
                 cpt = "Figure 3, Continued", ftnt = "")
```

```{r}
doc <- Table.Single(dt = out.fac.resid, cmp = So.res, cuz = cuz.top.row,
                 cpt = "Figure 4", ftnt = "Continued ...")
doc <- Table.Single(dt = out.fac.resid, cmp = So.res, cuz = cuz.bot.row,
                 cpt = "Figure 4, Continued", ftnt = "")
```

```{r}
out.pmc.race <- out.pmc.race %>% 
  mutate(Race = ifelse(Race == "A", "Asian",
                       ifelse(Race == "B", "Black",
                              ifelse(Race == "H", "Hispanic", "White"))))

doc <- Table.Multi.race(dt = out.pmc.race, cmp = cmp.multi, cuz = cuz.top.row,
                 cpt = "Figure 5", ftnt = "Continued ...")
doc <- Table.Multi.race(dt = out.pmc.race, cmp = cmp.multi, cuz = cuz.bot.row,
                 cpt = "Figure 5, Continued", ftnt = "")
```

```{r}
out.fac.race <- out.fac.race %>% 
  mutate(Race = ifelse(Race == "A", "Asian",
                       ifelse(Race == "B", "Black",
                              ifelse(Race == "H", "Hispanic", "White"))))

doc <- Table.Multi.race(dt = out.fac.race, cmp = So.multi, cuz = cuz.top.row,
                 cpt = "Figure 6", ftnt = "Continued ...")
doc <- Table.Multi.race(dt = out.fac.race, cmp = So.multi, cuz = cuz.bot.row,
                 cpt = "Figure 6, Continued", ftnt = "")
```

```{r}
out.pmc.age <- out.pmc.age %>% 
  mutate(Age = ifelse(Age == "LE75", "65-75", "> 75"),
         Age = factor(Age, levels = c("65-75", "> 75")))

doc <- Table.Multi.age(dt = out.pmc.age, cmp = cmp.multi, cuz = cuz.top.row,
                 cpt = "Figure 7", ftnt = "Continued ...")
doc <- Table.Multi.age(dt = out.pmc.age, cmp = cmp.multi, cuz = cuz.bot.row,
                 cpt = "Figure 7, Continued", ftnt = "")
```

```{r}
out.fac.age <- out.fac.age %>% 
  mutate(Age = ifelse(Age == "LE75", "65-75", "> 75"),
         Age = factor(Age, levels = c("65-75", "> 75")))

doc <- Table.Multi.age(dt = out.fac.age, cmp = So.multi, cuz = cuz.top.row,
                 cpt = "Figure 8", ftnt = "Continued ...")
doc <- Table.Multi.age(dt = out.fac.age, cmp = So.multi, cuz = cuz.bot.row,
                 cpt = "Figure 8, Continued", ftnt = "")
```

```{r}
out.pmc.sex <- out.pmc.sex %>% 
  mutate(Sex = ifelse(Sex == "F", "Female", "Male"))

doc <- Table.Multi.sex(dt = out.pmc.sex, cmp = cmp.multi, cuz = cuz.top.row,
                 cpt = "Figure 9", ftnt = "Continued ...")
doc <- Table.Multi.sex(dt = out.pmc.sex, cmp = cmp.multi, cuz = cuz.bot.row,
                 cpt = "Figure 9, Continued", ftnt = "")
```

```{r}
out.fac.sex <- out.fac.sex %>% 
  mutate(Sex = ifelse(Sex == "F", "Female", "Male"))

doc <- Table.Multi.sex(dt = out.fac.sex, cmp = So.multi, cuz = cuz.top.row,
                 cpt = "Figure 10", ftnt = "Continued ...")
doc <- Table.Multi.sex(dt = out.fac.sex, cmp = So.multi, cuz = cuz.bot.row,
                 cpt = "Figure 10, Continued", ftnt = "")
```

```{r}
doc <- Table.Single(dt = out.pmc.urbanicity %>% filter(Loc.Set == "urban") %>%
                      select(-Loc.Set),
                    cmp = cmp.main, cuz = cuz.top.row,
                 cpt = "Figure 11", ftnt = "Continued ...")
doc <- Table.Single(dt = out.pmc.urbanicity%>% filter(Loc.Set == "urban") %>% 
                      select(-Loc.Set),
                    cmp = cmp.main, cuz = cuz.bot.row,
                 cpt = "Figure 11, Continued", ftnt = "")
```

```{r}
doc <- Table.Single(dt = out.fac.urbanicity %>% filter(Loc.Set == "urban") %>% 
                      select(-Loc.Set),
                    cmp = So.main, cuz = cuz.top.row,
                 cpt = "Figure 12", ftnt = "Continued ...")
doc <- Table.Single(dt = out.fac.urbanicity%>% filter(Loc.Set == "urban") %>% 
                      select(-Loc.Set),
                    cmp = So.main, cuz = cuz.bot.row,
                 cpt = "Figure 12, Continued", ftnt = "")
```

```{r}
out.pmc.sescat <- out.pmc.sescat %>% 
  mutate(SES.Cat = ifelse(SES.Cat == "L", "Low",
                          ifelse(SES.Cat == "M", "Middle", "High")),
         SES.Cat = factor(SES.Cat, levels = c("Low", "Middle", "High")))

doc <- Table.Multi.sescat(dt = out.pmc.sescat, cmp = cmp.multi, cuz = cuz.top.row,
                 cpt = "Figure 13", ftnt = "Continued ...")
doc <- Table.Multi.sescat(dt = out.pmc.sescat, cmp = cmp.multi, cuz = cuz.bot.row,
                 cpt = "Figure 13, Continued", ftnt = "")
```

```{r}
out.pmc.region <- out.pmc.region %>% 
  mutate(Region = ifelse(Region == "west", "West",
                         ifelse(Region == "midwest", "Midwest",
                                ifelse(Region == "south", "South", "Northeast"))),
         Region = factor(Region, levels = c("West", "Midwest", "South", "Northeast")))

doc <- Table.Multi.region(dt = out.pmc.region, cmp = cmp.multi, cuz = cuz.top.row,
                 cpt = "Figure 14", ftnt = "Continued ...")
doc <- Table.Multi.region(dt = out.pmc.region, cmp = cmp.multi, cuz = cuz.bot.row,
                 cpt = "Figure 14, Continued", ftnt = "")
```

```{r}
out.fac.region <- out.fac.region %>% 
  mutate(Region = ifelse(Region == "west", "West",
                         ifelse(Region == "midwest", "Midwest",
                                ifelse(Region == "south", "South", "Northeast"))),
         Region = factor(Region, levels = c("West", "Midwest", "South", "Northeast")))

doc <- Table.Multi.region(dt = out.fac.region, cmp = So.multi, cuz = cuz.top.row,
                 cpt = "Figure 15", ftnt = "Continued ...")
doc <- Table.Multi.region(dt = out.fac.region, cmp = So.multi, cuz = cuz.bot.row,
                 cpt = "Figure 15, Continued", ftnt = "")
```

```{r}
out.pmc.resid.region <- out.pmc.resid.region %>% 
  mutate(Region = ifelse(Region == "west", "West",
                         ifelse(Region == "midwest", "Midwest",
                                ifelse(Region == "south", "South", "Northeast"))),
         Region = factor(Region, levels = c("West", "Midwest", "South", "Northeast")))

doc <- Table.Multi.region(dt = out.pmc.resid.region, cmp = cmp.multi.res, cuz = cuz.top.row,
                 cpt = "Figure 16", ftnt = "Continued ...")
doc <- Table.Multi.region(dt = out.pmc.resid.region, cmp = cmp.multi.res, cuz = cuz.bot.row,
                 cpt = "Figure 16, Continued", ftnt = "")
```

```{r}
out.pmc.brfss <- out.pmc.brfss %>% 
  mutate(BRFSS = ifelse(BRFSS == "brfss0", "Base", "BRFSS-Adjusted"))

doc <- Table.Multi.brfss(dt = out.pmc.brfss, cmp = cmp.multi, cuz = cuz.top.row,
                 cpt = "Figure 17", ftnt = "Continued ...")
doc <- Table.Multi.brfss(dt = out.pmc.brfss, cmp = cmp.multi, cuz = cuz.bot.row,
                 cpt = "Figure 17, Continued", ftnt = "")
```

```{r}
out.pmc.buffer <- out.pmc.buffer %>% 
  mutate(Buffer = ifelse(Buffer == "6", "6-miles",
                         ifelse(Buffer == "12", "12-miles", "24-miles")),
         Buffer = factor(Buffer, levels = c("6-miles", "12-miles", "24-miles")))

doc <- Table.Multi.buffer(dt = out.pmc.buffer, cmp = cmp.multi, cuz = cuz.top.row,
                 cpt = "Figure 18", ftnt = "Continued ...")
doc <- Table.Multi.buffer(dt = out.pmc.buffer, cmp = cmp.multi, cuz = cuz.bot.row,
                 cpt = "Figure 18, Continued", ftnt = "")
```

```{r}
out.pmc.expo <- out.pmc.expo %>% 
  mutate(Window = ifelse(Window == "1", "1-yr",
                         ifelse(Window == "2", "2-yr",
                                ifelse(Window == "3", "3-yr",
                                       ifelse(Window == "4", "4-yr", "5-yr")))))

doc <- Table.Multi.expo(dt = out.pmc.expo, cmp = cmp.multi, cuz = cuz.top.row,
                 cpt = "Figure 19", ftnt = "Continued ...")
doc <- Table.Multi.expo(dt = out.pmc.expo, cmp = cmp.multi, cuz = cuz.bot.row,
                 cpt = "Figure 19, Continued", ftnt = "")
```

```{r}
print(doc, target = here('output','table','supplement-tables.docx'))
```